---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# bacphene

<!-- badges: start -->
<!-- badges: end -->

The goal of bacphene is to import bacterial phenotype data from bacdive.org into R.

## Installation

You can install bacphene from [github](https://github.com/scottdaniel/bacphene) with:

``` r
library(devtools)
install_github(repo = "scottdaniel/bacphene")
```

After this, you MUST register at bacdive.org [here](https://bacdive.dsmz.de/api/bacdive/registration/register/). Otherwise, you will not be able to access the API and get phenotype data. 

## Example

This will only work once you have gotten an account at bacdive.org:

```

library(bacphene)

user <- "usually_your_email"
passwd <- "a_good_password"

strain_list <- getStrains(page = 1, 
  genus = 'Bacteroides', 
  species = 'xylanisolvens', 
  userpassword = paste0(user,':',passwd))
  
phenotypes_list <- getStrainData(strain_list,
    selection=1,
    userpassword = paste0(user,':',passwd))

```

However, to facilitate demonstration, I have run the previous codeblock and written the results lists to flat files here:
./examples/Bacteroides_xylanisolvens.RData

Here we can see that there are (as of 10/19/2020) three strains in the bacdive.org database with the species designation of Bacteroides xylanisolvens:


```{r}

load("./examples/Bacteroides_xylanisolvens.RData")

strain_list$results

```

Selecting the first strain (which you can set in [getStrainData](https://github.com/scottdaniel/bacphene/blob/master/R/bacPhene.R) with 'selection=1') we get the following information:

```{r}

names(phenotypes_list)

```

