---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# bacphene

<!-- badges: start -->
<!-- badges: end -->

The goal of bacphene is to generate a table of phenotype characteristics about your samples. Input is a table of Samples with counts per taxa. 

The package utilizes the official [BacDive R package] that accesses
bacterial phenotype data from [bacdive.org].

## Installation

You can install bacphene from the [github repo]. You also need to install the BacDive R package (if you want the full data - see below).

``` r
library(devtools)
install_github(repo = "scottdaniel/bacphene")
install.packages("BacDive", repos="http://R-Forge.R-project.org")
```

After this, you may register at bacdive.org [here](https://api.bacdive.dsmz.de/login). 

If you do not register, you will not be able to download the full information on strains in BacDive. Instead, what you will get is access to three data-frames: `bacdive_phenotypes`, `bacdive_susceptibility`, and `bacdive_enzymes`. Which look like this:

```{r}
library(bacphene)
head(bacdive_phenotypes)
```

```{r}
head(bacdive_susceptibility)
```

```{r}
head(bacdive_enzymes)
```


### Recommended

Edit your $HOME/.Renviron file (you can open in R with `usethis::edit_r_environ()`) and add your bacdive credentials like so:
```
DSMZ_API_USER=your_email@something.com
DSMZ_API_PASSWORD=your_password
```

# References

When using BacDive for research please consider citing the following paper:

BacDive in 2019: bacterial phenotypic data for High-throughput biodiversity analysis.
Reimer, L. C., Vetcininova, A., Sardà Carbasse, J., Söhngen, C., Gleim, D., Ebeling, C., Overmann, J.
[Nucleic Acids Research; database issue 2019](https://academic.oup.com/nar/article/47/D1/D631/5106998).

To cite this package enter `citation("bacphene")` in your R console.

[BacDive R package]: https://r-forge.r-project.org/R/?group_id=1573
[bacdive.org]: https://bacdive.dsmz.de
[github repo]: https://github.com/scottdaniel/bacphene
