---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# bacphene

<!-- badges: start -->
<!-- badges: end -->

The goal of bacphene is to generate a table of phenotype characteristics about your samples. Input is a table of Samples with counts per taxa. 

The package utilizes the official [BacDive R package](https://r-forge.r-project.org/R/?group_id=1573) that accesses
bacterial phenotype data from [bacdive.org](https://bacdive.dsmz.de/).

## Installation

You can install bacphene from [github](https://github.com/scottdaniel/bacphene). You also need to install the BacDive R package (if you want the full data - see below).

``` r
library(devtools)
install_github(repo = "scottdaniel/bacphene")
install.packages("BacDive", repos="http://R-Forge.R-project.org")
```

After this, you may register at bacdive.org [here](https://api.bacdive.dsmz.de/login). 

If you do not register, you will not be able to download the full information on strains in BacDive. Instead, what you will get is access to two data-frames: `bacdive_phenotypes` and `bacdive_susceptibility`. Which look like this:

```{r}
library(bacphene)
head(bacdive_phenotypes)
```

```{r}
head(bacdive_susceptibility)
```

* Recommended *

Edit your $HOME/.Renviron file (you can open in R with `usethis::edit_r_environ()`) and add your bacdive credentials like so:
```
DSMZ_API_USER=your_email@something.com
DSMZ_API_PASSWORD=your_password
```
